---
title: "Degree_day_linear_models_V2"
author: "Sara DeLaurentis"
date: "3/23/2023"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## LIBRARIES

library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(dplyr)
library(glue)
library(lme4)
library(lmerTest)
library(emmeans)
library(kableExtra)
library(Hmisc)


```

Purpose: Analysis of Degree Day data


## Data preparation ##

#### Data read-in


```{r read.in, echo = TRUE}
data <- read_csv("../ALL_A.csv")
data$`...1` = NULL #remove strange column that was created upon read-in.
head(data)

data <- data[!(data$site == "D5B"),] #remove all D5B sensors.
D5B.chk <- data[data$site == "D5B",]
D5B.chk

```



#### Summarize mean temps for each date

Create a new df:
 mutate date.time to POSIXct object, then summarize temperature measurements into one meantemp measurement for each date.
 
 
#### Rename the grouping column to "date", from the automatically created 'date(date.time)'


```{r summarize.daily, echo = TRUE}

dailymeans.df <- mutate(data, date.time = as.POSIXct(date.time, format = "%Y-%m-%d %H:%M")) %>% 
  group_by(name, date(date.time)) %>% 
  summarise(meantemp = mean(value))
colnames(dailymeans.df)[2]<- "date" #rename grouping column

#head(dailymeans.df)

```


#### Problem-solving

```{r problem.solve, echo = TRUE}

C2A_R2_m10.means <- dailymeans.df %>% filter(grepl("C2A_R2_m10",name) == TRUE)
#view(C2A_R2_m10.means)

C2A_R3_m10.means <- dailymeans.df %>% filter(grepl("C2A_R3_m10", name) == TRUE)
#view(C2A_R3_m10.means)

```

#### Remove all values less than or equal to 0

```{r filter.zero, echo = TRUE}

dailymeans.df <- dailymeans.df %>% filter(meantemp > 0)

```

#### code for remembering (hidden)

```{r remembering, echo = FALSE}
#DD.df <- dailymeans.df %>% 
#  group_by(SRP.name) %>% 
#  summarise(SRP.name = SRP.name,
#            degree.days = sum(meantemp)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
#  distinct(SRP.name, .keep_all = TRUE) %>% 
#  ungroup()
#view(DD.df)  


#Helpful code for remembering.

#  C2A_R1_m10_dailymean <- mutate(C2A_R1_m10,'date.time' = mdy_hm(date.time)) %>% 
#  separate('date.time',
#           into = c('longdate', 'time'),
#           sep = ' ') %>% 
#  separate('longdate',
#           into = c('year', 'month', 'day'),
#           sep = '-',
#           remove = FALSE) %>% 
#  group_by(year, month, day) %>% 
#  summarise(meantemp = mean(value)) 
```



### TIME BREAKS

```{r TIME BREAKS, echo = TRUE}

time.start <- as.Date("2020-04-01")
#time.break.one <-
#time.break.two <- 
time.end <- as.Date("2020-08-27")
print(time.start)
print(time.end)


## subsetting the dataframe using the time breaks.
#Selects data that is greater than time.start AND less than time.end
DD.df.cut <- dailymeans.df[(dailymeans.df$date >= time.start) & (dailymeans.df$date <= time.end),]

#view(DD.df.cut)

```


#### Add in metadata columns for final dataframe

```{r attach meta, echo = TRUE}

#Split the name column into 5 separate columns
meta.df <- str_split_fixed(DD.df.cut$name, "_", n = 5)
#head(meta.df)

# Name and attach the columns from meta.df
DD.df.cut <- DD.df.cut %>% add_column(site = meta.df[,1], rep = meta.df[,2], position = meta.df[,3], buttonID = meta.df[,4], season = meta.df[,5],
                         .before = "date")
#head(DD.df.cut)

```


#### Cutting out incomplete datasets. Removes the group if the length is less than 50 days.

```{r rm.incomplete, echo = TRUE}

DD.df.cut <- DD.df.cut %>% group_by(name) %>% filter((length(name) > 50) == TRUE)
#view(DD.df.cut)

```
     
#### Summarise, or reframe. Final step for response variable df 

Group by site, rep, and position, and sum the mean temperatures in the cut dataframe.
This gives you the response variable: degree days tabulated over the desired timeframe.

```{r merge complete table, echo = TRUE}

DDsums.df.pre <- DD.df.cut %>% 
    group_by(site, rep, position) %>% 
    reframe(site = site, position = position,
              degree.days = sum(meantemp)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
    distinct(site, rep, position, .keep_all = TRUE) #%>% 
    #ungroup()
  #view(DDsums.df.pre)


## Creating full (expected) dataframe, to merge with the current dataframe.  

#  site <- c(rep("C2A", 14), rep("C2B", 14), rep("C5A", 14), rep("C5B", 14),
#          rep("D2A", 14), rep("D2B", 14), rep("D5A", 14), rep("D5B", 14))
#print(site)

#rep <- c("R0", rep("R1",5), rep("R2", 4), rep("R3",4))  %>% 
#  rep(8)
#print(rep)

sensor.list <- as_tibble(read_csv("../sensor_list.csv"))
#view(sensor.list)


#Add column to DDsums.df and sensor.list to give a unique column to merge with.
DDsums.df.A <- DDsums.df.pre %>% 
  mutate(srp.name = paste(site, rep, position, sep = "_"))
#head(DDsums.df.A)

sensor.list.A <- sensor.list %>% 
  mutate(srp.name = paste(site, rep, position, sep = "_"))
#head(sensor.list.A)

#DDsums.df.new <- merge(x = sensor.list.A, y = DDsums.df.A, by = "srp.name", all.x = TRUE)
DDsums.df <- merge(x = sensor.list.A, y = DDsums.df.A, all.x = TRUE)

 
#view(DDsums.df)

```



#### Adding treatment columns:
  
Worm invasion level (worm_lvl, LOW or HIGH)

```{r trt cols, echo = TRUE}

DDsums.df.a <- DDsums.df %>% mutate(worm_lvl = if_else(grepl("2", DDsums.df$site), "LOW", "HIGH"))
#head(DDsums.df.a)
#view(DDsums.df.a) 

#Vegetation (Veg, Coniferous or Deciduous)

DDsums.df.b <- DDsums.df.a %>% mutate(Veg = if_else(grepl("C", DDsums.df$site), "Coniferous", "Deciduous"))
#head(DDsums.df.b)
#view(DDsums.df.b)  

DDSUMS.df <- DDsums.df.b # rename df
#head(DDSUMS.df)

```

## Bring in and attach OA horizon data ##

 See O_horizon_prelim.R to view summary statistics


```{r OA data}

OA.dat <- read.csv("../OA_thickness_2019_21_clean_20230325.csv")

# set classes for columns
OA.dat$inv_lvl <- as.character(OA.dat$inv_lvl)
OA.dat$Oi_thk <- as.numeric(OA.dat$Oi_thk)
OA.dat$Oe_thk <- as.numeric(OA.dat$Oe_thk)
OA.dat$Oa_thk <- as.numeric(OA.dat$Oa_thk)
OA.dat$O.A_thk <- as.numeric(OA.dat$O.A_thk)
OA.dat$bulk_density <- as.numeric(OA.dat$bulk_density)

# extract mean O thickness, attach to DD data

OA.means.df <- OA.dat %>% 
  filter(!is.na(totalO_thk)) %>% 
  group_by(site) %>% 
  summarise(Omean = round(mean(totalO_thk, na.rm = TRUE), digits = 2))
OA.means.df
# attach
DDSUMS.ALL <- DDSUMS.df %>% 
  left_join(OA.means.df)
head(DDSUMS.ALL)  

#
#df["Column Name"][df["Column Name"] == "Old Value"] <- "New Value"
DDSUMS.ALL["position"][DDSUMS.ALL["position"] == "m01surf"] <- "lsurf"
DDSUMS.ALL["position"][DDSUMS.ALL["position"] == "m02surf"] <- "m0surf"


```


## SIMPLE TESTS - one point per site ##


```{r Simple.df}

simple.df <- DDSUMS.ALL %>% 
  group_by(site, position) %>%
  reframe(site = site, 
          position = position, 
          worm_lvl = worm_lvl, 
          Veg = Veg, 
          Omean = Omean, 
          degree.days = mean(degree.days, na.rm = TRUE)) %>% 
  distinct(site, position, .keep_all = TRUE)
  
simple.df <- simple.df %>% mutate(treatment = substr(site,1,2))


air.df <- simple.df[simple.df$position == "air", c("site", "degree.days")]
colnames(air.df)[2] = "air.DD"
view(air.df)

simple.df <- simple.df %>% 
  left_join(air.df)
#DDsums.df.pre <- DD.df.cut %>% 
#    group_by(site, rep, position) %>% 
#    reframe(site = site, rep = rep, position = position, worm_lvl = worm_lvl, Veg = Veg, Omean = Omean,
#              degree.days = sum(meantemp)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
#    distinct(site, rep, position, .keep_all = TRUE) #%>% 
    #ungroup()
  #view(DDsums.df.pre) 

view(simple.df)

```

```{r simple tests}

# lsurf

#generate lsurf df for analysis
DDSUMS.lsurf <- simple.df %>% filter(position == "lsurf" | position == "m01surf")
view(DDSUMS.lsurf)

fm1.lsurf <- lm(degree.days ~ treatment, data = DDSUMS.lsurf)
anova(fm1.lsurf)

# m0surf

#generate m0surf df for analysis
DDSUMS.m0surf <- simple.df %>% filter(position == "m0surf")
view(DDSUMS.m0surf)

fm1.m0surf <- lm(degree.days ~ treatment, data = DDSUMS.m0surf)
anova(fm1.m0surf)
summary(fm1.m0surf)

# m10

#generate m10 df for analysis
DDSUMS.m10 <- simple.df %>% filter(position == "m10")
view(DDSUMS.m10)

fm1.m10 <- lm(degree.days ~ treatment, data = DDSUMS.m10)
anova(fm1.m10)

# m30

#generate m30 df for analysis
DDSUMS.m30 <- simple.df %>% filter(position == "m30" | position == "m01surf")
view(DDSUMS.m30)

fm1.m30 <- lm(degree.days ~ treatment, data = DDSUMS.m30)
anova(fm1.m30)

```

```{r linear regression}

# lsurf

#generate lsurf df for analysis
DDSUMS.lsurf <- simple.df %>% filter(position == "lsurf" | position == "m01surf")
view(DDSUMS.lsurf)

fm.Othk.lsurf <- lm(degree.days ~ Veg + Omean, data = DDSUMS.lsurf)
anova(fm.Othk.lsurf)
summary(fm.Othk.lsurf)

# m0surf

#generate m0surf df for analysis
DDSUMS.m0surf <- simple.df %>% filter(position == "m0surf")
view(DDSUMS.m0surf)

fm.Othk.m0surf <- lm(degree.days ~ Veg + Omean, data = DDSUMS.m0surf)
anova(fm.Othk.m0surf)
summary(fm.Othk.m0surf)

# m10

#generate m10 df for analysis
DDSUMS.m10 <- simple.df %>% filter(position == "m10")
view(DDSUMS.m10)

fm.Othk.m10 <- lm(degree.days ~ Veg + Omean, data = DDSUMS.m10)
anova(fm.Othk.m10)
summary(fm.Othk.m10)
 
# m30

#generate m30 df for analysis
DDSUMS.m30 <- simple.df %>% filter(position == "m30" | position == "m01surf")
view(DDSUMS.m30)

fm.Othk.m30 <- lm(degree.days ~ Veg + Omean, data = DDSUMS.m30)
anova(fm.Othk.m30)
summary(fm.Othk.m30)


```

## LINEAR MODEL - replicates are separate points ##



#### DD vegxworm lsurf mixed (DDSUMS.lsurf)

#This model has degree.days as a function of vegetation (Veg) and worm invasion intensity (worm_lvl), as well as the interaction as slopes, with site as a random effect.
#PARAMETERS: Timeframe: 2020-04-01 - 2020-09-30, position: lsurf


```{r mod1 lsurf}

#generate lsurf df for analysis
DDSUMS.lsurf <- DDSUMS.ALL %>% filter(position == "lsurf" | position == "m01surf")
view(DDSUMS.lsurf)

mod1.lsurf <- lmer(data = DDSUMS.lsurf, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl + (1|site))

#View object
mod1.lsurf
summary(mod1.lsurf)
plot(residuals(mod1.lsurf) ∼ fitted.values(mod1.lsurf))

#Error:
#" boundary (singular) fit: see help('isSingular') "
#2023-03-16: After filling in data gaps with NAs, no more singular fit error.

```

#### Creation of mod2: What happens if we remove the random effect for site?

lm DD vegxworm lsurf (mod2.lsurf)
linear model with no random effect:
PARAMETERS: timeframe = 2020-04-01 - 2020-09-30, position: lsurf

```{r mod2 lsurf}

mod2.lsurf <- lm(data = DDSUMS.lsurf, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl)
summary(mod2.lsurf)
#non-significant slopes for lsurf, which makes sense

plot(residuals(mod2.lsurf) ∼ fitted.values(mod2.lsurf))

```


#### m0surf models

```{r mod1.2 m0surf}
#creation of m0surf object
DDSUMS.m0surf <- DDSUMS.ALL %>% filter(position == "m0surf" | position == "m02surf")
view(DDSUMS.m0surf)

##m0surf lmer
mod1.m0surf <- lmer(data = DDSUMS.m0surf, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl + (1|site))
summary(mod1.m0surf)
plot(residuals(mod1.m0surf) ∼ fitted.values(mod1.m0surf))

##m0surf lm
mod2.m0surf <- lm(data = DDSUMS.m0surf, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl)
summary(mod2.m0surf)
plot(residuals(mod2.m0surf) ∼ fitted.values(mod2.m0surf))
plot(mod2.m0surf)

```

#### m10 models

#### mixed-effects DD vegxworm m10
#### lmer for m10. 
Note: for the regular linear models, this is not taking into account the natural variation by site, that may be due to soil type and solar radiation, slope, etc.
PARAMETERS: timeframe = 2020-04-01 - 2020-09-30, position: m10

```{r mod 1.2 m10}
##m10 models

#create m10 object:
DDSUMS.m10 <- DDSUMS.ALL %>% filter(position == "m10")
print(DDSUMS.m10)
#view(DDSUMS.m10)

##m10 lmer
mod1.m10 <- lmer(data = DDSUMS.m10, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl + (1|site))
summary(mod1.m10)

##m10 lm
mod2.m10 <- lm(data = DDSUMS.m10, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl)
summary(mod2.m10)

```
Significant effect for worm_lvl, no significant effect for vegetation type, or interaction.



####m30 models

```{r mod 1.2 m30}

#create m30 object:
DDSUMS.m30 <- DDSUMS.ALL %>% filter(position == "m30")
print(DDSUMS.m30)
#view(DDSUMS.m30)

##m30 lmer
mod1.m30 <- lmer(data = DDSUMS.m30, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl + (1|site))
summary(mod1.m30)

##m30 lm
mod2.m30 <- lm(data = DDSUMS.m30, formula = degree.days ~ Veg + worm_lvl + Veg*worm_lvl)
summary(mod2.m30)

```


#### air sensor DD data

```{r air}
#create air object:
DDSUMS.air <- DDSUMS.ALL %>% filter(position == "air")
print(DDSUMS.air)

#summary statistics for air DD
#create df for air data at each site:
air.DD.df <- data_frame(site = DDSUMS.air$site, degree.days = DDSUMS.air$degree.days)
print(air.DD.df)
print(summary(DDSUMS.air$degree.days))

```

## Model including O horizon data ##

#### Creation of mod3: Including mean O horizon thickness



```{r mod3.4 lsurf}

#create lsurf object:
DDSUMS.lsurf <- DDSUMS.ALL %>% filter(position == "lsurf")
print(DDSUMS.lsurf)
#view(DDSUMS.lsurf)

##lsurf lmer
mod3.lsurf <- lmer(data = DDSUMS.lsurf, formula = degree.days ~ Veg + Omean + Veg * Omean + (1|site))
summary(mod3.lsurf)

##lsurf lm
mod4.lsurf <- lm(data = DDSUMS.lsurf, formula = degree.days ~ Veg + Omean + Veg * Omean)
summary(mod4.lsurf)
plot(mod4.lsurf)

d<-density(mod4.lsurf[['residuals']])
plot(d,main='Residual KDE Plot',xlab='Residual value')

```



```{r mod3.4 m0surf}

#creation of m0surf object
DDSUMS.m0surf <- DDSUMS.ALL %>% filter(position == "m0surf" | position == "m02surf")
#view(DDSUMS.m0surf)

##m0surf lmer
mod3.m0surf <- lmer(data = DDSUMS.m0surf, formula = degree.days ~ Veg + Omean + Veg * Omean + (1|site))
summary(mod3.m0surf)

##m0surf lm
mod4.m0surf <- lm(data = DDSUMS.m0surf, formula = degree.days ~ Veg + Omean + Veg * Omean)
summary(mod4.m0surf)
plot(mod4.m0surf)

```


```{r mod3.4 m10}

##m10 models

#create m10 object:
DDSUMS.m10 <- DDSUMS.ALL %>% filter(position == "m10")
#print(DDSUMS.m10)
#view(DDSUMS.m10)

##m10 lmer
mod3.m10 <- lmer(data = DDSUMS.m10, formula = degree.days ~ Veg + Omean + Veg * Omean + (1|site))
summary(mod3.m10)

##m10 lm
mod4.m10 <- lm(data = DDSUMS.m10, formula = degree.days ~ Veg + Omean + Veg * Omean)
summary(mod4.m10)

```



```{r mod3.4 m30}

#create m30 object:
DDSUMS.m30 <- DDSUMS.ALL %>% filter(position == "m30")
print(DDSUMS.m30)
#view(DDSUMS.m30)

##m30 lmer
mod3.m30 <- lmer(data = DDSUMS.m30, formula = degree.days ~ Veg + Omean + Veg * Omean + (1|site))
summary(mod3.m30)

##m30 lm
mod4.m30 <- lm(data = DDSUMS.m30, formula = degree.days ~ Veg + Omean + Veg * Omean)
summary(mod4.m30)

```

#### Subtract replicate measurements from Air temp at sites with AT.

