---
title: "Degree_days_for_Amynthas"
author: "Sara DeLaurentis"
date: "4/10/2024"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## LIBRARIES

library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(dplyr)
library(glue)
library(lme4)
library(lmerTest)
library(emmeans)
library(kableExtra)
library(Hmisc)


```

Purpose: Analysis of Degree Day data


## Data preparation ##

#### Data read-in


```{r read.in, echo = TRUE}
data <- read_csv("../ALL_A.csv")
data$`...1` = NULL #remove strange column that was created upon read-in.
head(data)

data <- data[!(data$site == "D5B"),] #remove all D5B sensors.
D5B.chk <- data[data$site == "D5B",]
D5B.chk

```

### Cut out sensors we don't want or need

```{r sensors wanted}

DD.wanted <- data %>% filter(position == "air" | position == "m0surf" | position == "m01surf" | position == "m02surf" | position == "m10")

head(DD.wanted)

```

#### Summarize mean temps for each date

Create a new df:
 mutate date.time to POSIXct object, then summarize temperature measurements into one meantemp measurement for each date.
 
 
#### Rename the grouping column to "date", from the automatically created 'date(date.time)'


```{r summarize.daily, echo = TRUE}

dailymeans.df <- mutate(DD.wanted, date.time = as.POSIXct(date.time, format = "%Y-%m-%d %H:%M")) %>% 
  group_by(name, date(date.time)) %>% 
  summarise(meantemp = mean(value))
colnames(dailymeans.df)[2]<- "date" #rename grouping column

head(dailymeans.df)

```


#### Problem-solving

```{r problem.solve, echo = TRUE}

C2A_R2_m10.means <- dailymeans.df %>% filter(grepl("C2A_R2_m10",name) == TRUE)
#view(C2A_R2_m10.means)

C2A_R3_m10.means <- dailymeans.df %>% filter(grepl("C2A_R3_m10", name) == TRUE)
#view(C2A_R3_m10.means)

```


#### code for remembering (hidden)

```{r remembering, echo = FALSE}
#DD.df <- dailymeans.df %>% 
#  group_by(SRP.name) %>% 
#  summarise(SRP.name = SRP.name,
#            degree.days = sum(meantemp)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
#  distinct(SRP.name, .keep_all = TRUE) %>% 
#  ungroup()
#view(DD.df)  


#Helpful code for remembering.

#  C2A_R1_m10_dailymean <- mutate(C2A_R1_m10,'date.time' = mdy_hm(date.time)) %>% 
#  separate('date.time',
#           into = c('longdate', 'time'),
#           sep = ' ') %>% 
#  separate('longdate',
#           into = c('year', 'month', 'day'),
#           sep = '-',
#           remove = FALSE) %>% 
#  group_by(year, month, day) %>% 
#  summarise(meantemp = mean(value)) 
```



### TIME BREAKS

```{r TIME BREAKS, echo = TRUE}

#time.start <- as.Date("2021-04-01")
#time.break.one <-
#time.break.two <- 
#time.end <- as.Date("2021-08-27")


time.start <- as.Date("2020-10-01")
#time.break.one <-
#time.break.two <- 
time.end <- as.Date("2021-09-30")

print(time.start)
print(time.end)
## subsetting the dataframe using the time breaks.
#Selects data that is greater than time.start AND less than time.end
DD.df.cut <- dailymeans.df[(dailymeans.df$date >= time.start) & (dailymeans.df$date <= time.end),]

#view(DD.df.cut)

```

#### Cutting out incomplete datasets. Removes the group if the length is less than 50 days.

```{r rm.incomplete, echo = TRUE}

DD.df.cut <- DD.df.cut %>% group_by(name) %>% filter((length(name) > 50) == TRUE)
view(DD.df.cut)


```

#### Remove all values less than or equal to 0, create new column with base temperature subtracted.

```{r filter.zero, echo = TRUE}
GDD5.df <- DD.df.cut
GDD5.df$GDD5 <- GDD5.df$meantemp - 5  #create new column by subtracting 5 from meantemp.
GDD5.df <- GDD5.df %>% filter(GDD5 > 0)

head(GDD5.df)

```
#### Add in metadata columns for final dataframe

```{r attach meta, echo = TRUE}

#Split the name column into 5 separate columns
meta.df <- str_split_fixed(GDD5.df$name, "_", n = 5)
#head(meta.df)

# Name and attach the columns from meta.df
GDD5.df <- GDD5.df %>% add_column(site = meta.df[,1], rep = meta.df[,2], position = meta.df[,3], buttonID = meta.df[,4], season = meta.df[,5],
                         .before = "date")
head(GDD5.df)

```

#### Summarise, or reframe. Final step for response variable df 

Group by site, rep, and position, and sum the mean temperatures in the cut dataframe.
This gives you the response variable: degree days tabulated over the desired timeframe.

```{r merge complete table, echo = TRUE}

DDsums.df.pre <- GDD5.df %>% 
  group_by(site, rep, position, season) %>% 
  reframe(site = site, position = position,
          degree.days = sum(GDD5)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
  distinct(site, rep, position, season, .keep_all = TRUE) #%>% 
#ungroup()
view(DDsums.df.pre)

## Creating full (expected) dataframe, to merge with the current dataframe.  

#  site <- c(rep("C2A", 14), rep("C2B", 14), rep("C5A", 14), rep("C5B", 14),
#          rep("D2A", 14), rep("D2B", 14), rep("D5A", 14), rep("D5B", 14))
#print(site)

#rep <- c("R0", rep("R1",5), rep("R2", 4), rep("R3",4))  %>% 
#  rep(8)
#print(rep)

sensor.list <- as_tibble(read_csv("../sensor_list.csv"))
#view(sensor.list)

#Add column to DDsums.df and sensor.list to give a unique column to merge with.
DDsums.df.A <- DDsums.df.pre %>% 
  mutate(srps.name = paste(site, rep, position, season, sep = "_"))
#head(DDsums.df.A)

sensor.list.A <- sensor.list %>% 
  mutate(srp.name = paste(site, rep, position, sep = "_"))
#head(sensor.list.A)

#DDsums.df.new <- merge(x = sensor.list.A, y = DDsums.df.A, by = "srp.name", all.x = TRUE)
DDsums.df <- merge(x = sensor.list.A, y = DDsums.df.A, all.x = TRUE)


view(DDsums.df)



```

#### Replace m01surf and m02surf labels

```{r remove m01surf and m02surf}

#format:
#df[df == "Old Value"] <- "New Value"

DDsums.df[DDsums.df == "m01surf"] <- "lsurf"
DDsums.df[DDsums.df == "m02surf"] <- "m0surf"

#print(DDsums.df$position == "m01surf" | DDsums.df$position == "m02surf")

```

#### Adding treatment columns:
  
Worm invasion level (worm_lvl, LOW or HIGH)

```{r trt cols, echo = TRUE}

DDsums.df.a <- DDsums.df %>% mutate(worm_lvl = if_else(grepl("2", DDsums.df$site), "LOW", "HIGH"))
#head(DDsums.df.a)
#view(DDsums.df.a) 

#Vegetation (Veg, Coniferous or Deciduous)

DDsums.df.b <- DDsums.df.a %>% mutate(Veg = if_else(grepl("C", DDsums.df$site), "Coniferous", "Deciduous"))
#head(DDsums.df.b)
#view(DDsums.df.b)  

DDSUMS.df <- DDsums.df.b # rename df
#head(DDSUMS.df)

```



