---
title: "Degree_days_for_Amynthas"
author: "Sara DeLaurentis"
date: "4/10/2024"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
    theme: cerulean
    highlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## LIBRARIES

library(tidyverse)
library(lubridate)
library(stringr)
library(broom)
library(dplyr)
library(glue)
library(lme4)
library(lmerTest)
library(emmeans)
library(kableExtra)
library(Hmisc)
library(zoo)


```

Purpose: Analysis of Degree Day data


## Data preparation ##

#### Data read-in


```{r read.in, echo = TRUE}
data <- read_csv("../ALL_A.csv")
data$`...1` = NULL #remove strange column that was created upon read-in.
head(data)

data <- data[!(data$site == "D5B"),] #remove all D5B sensors.
D5B.chk <- data[data$site == "D5B",]
D5B.chk

```

### Cut out sensors we don't want or need

```{r sensors wanted}

DD.wanted <- data %>% filter(position == "air" | position == "m0surf" | position == "m01surf" | position == "m02surf" | position == "m10")

head(DD.wanted)

```

#### Summarize mean temps for each date

Create a new df:
 mutate date.time to POSIXct object, then summarize temperature measurements into one meantemp measurement for each date.
 
 
#### Rename 

```{r }

dailymeans.df <- mutate(DD.wanted, date.time = as.POSIXct(date.time, format = "%Y-%m-%d %H:%M")) %>% 
  group_by(name, date(date.time)) %>% 
  summarise(meantemp = mean(value), GDD5 = (mean(value) -5))
colnames(dailymeans.df)[2]<- "date" #rename grouping column

#### Set all values less than 0 to 0.

dailymeans.df$GDD5 <- pmax(dailymeans.df$GDD5, 0)

head(dailymeans.df)
#the grouping column to "date", from the automatically created 'date(date.time)'

```

```{r summarize.daily, echo = TRUE}


```



### TIME BREAKS

Choose desired time.start / time.end option.

```{r TIME BREAKS, echo = TRUE}

#time.start <- as.Date("2021-04-01")
#time.break.one <- 
#time.break.two <- 
#time.end <- as.Date("2021-08-27")

time.start <- as.Date("2020-02-01")
time.end <- as.Date("2020-11-01")

#time.start <- as.Date("2021-02-01")
#time.end <- as.Date("2021-11-01")

#time.start <- as.Date("2020-05-13")
#time.end <- as.Date("2020-11-01")

#time.start <- as.Date("2021-05-13")
#time.end <- as.Date("2021-11-01")

#time.break.one <-
#time.break.two <- 

#time.start <- as.Date("2020-10-01")
#time.end <- as.Date("2021-09-30")


print(time.start)
print(time.end)
## subsetting the dataframe using the time breaks.
#Selects data that is greater than time.start AND less than time.end
DD.df.cut <- dailymeans.df[(dailymeans.df$date >= time.start) & (dailymeans.df$date <= time.end),]


#view(DD.df.cut)

```

#### Cutting out incomplete datasets. Removes the group if the length is less than 50 days.

```{r rm.incomplete, echo = TRUE}

DD.df.cut <- DD.df.cut %>% group_by(name) %>% filter((length(name) > 50) == TRUE)
#view(DD.df.cut)


```


#### Add in metadata columns for final dataframe

```{r attach meta, echo = TRUE}

#Split the name column into 5 separate columns
meta.df <- str_split_fixed(DD.df.cut$name, "_", n = 5)
#head(meta.df)

# Name and attach the columns from meta.df
DD.df.cut <- DD.df.cut %>% add_column(site = meta.df[,1], rep = meta.df[,2], position = meta.df[,3], buttonID = meta.df[,4], season = meta.df[,5],
                         .before = "date")

```


#### Cutting out incomplete datasets. Removes the group if the length is less than 50 days.

```{r rm.incomplete, echo = TRUE}

DD.df.cut <- DD.df.cut %>% group_by(name) %>% filter((length(name) > 50) == TRUE)
#view(DD.df.cut)

```
     
#### Summarise, or reframe. Final step for response variable df 

Group by site, rep, and position, and sum the mean temperatures in the cut dataframe.
This gives you the response variable: degree days tabulated over the desired timeframe.

#```{r merge complete table, echo = TRUE}

DDsums.df.pre <- DD.df.cut %>% 
  group_by(site, rep, position, season) %>% 
  reframe(site = site, position = position,
          degree.days = sum(meantemp)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
head(DD.df.cut)

#```

#### Summarise, or reframe. Final step for response variable df 

Group by site, rep, and position, and sum the mean temperatures in the cut dataframe.
This gives you the response variable: degree days tabulated over the desired timeframe.

## Rolling sums

```{r rollsum}

GDD5.sums.df <- DD.df.cut %>% 
  group_by(name) %>% 
  mutate(GDD5.sum = cumsum(GDD5))

head(GDD5.sums.df)
view(GDD5.sums.df)


test.df <- GDD5.sums.df %>% 
  group_by(name) %>% 
  filter(GDD5.sum >= 1000) %>% #removal of incomplete data sets.
  slice(1) %>% # takes the first occurrence if there is a tie
  ungroup()

view(test.df)

view(test.df[test.df$position == "lsurf" | test.df$position == "m0surf" | test.df$position == "m01surf" | test.df$position == "m02surf",])

view(test.df[test.df$position == "m10",])

print(min(test.df$date))
print(max(test.df$date))

```


#```{r merge complete table, echo = TRUE}

DDsums.df.pre <- test.df %>% 
  group_by(site, rep, position, season) %>% 
  reframe(site = site, position = position,
          degree.days = sum(GDD5)) %>%  #create degree.days column by outputting the sum of meantemp for each position. (Not sensor, but position)
  distinct(site, rep, position, season, .keep_all = TRUE) #%>% 
#ungroup()
view(DDsums.df.pre)

## Creating full (expected) dataframe, to merge with the current dataframe.  

#  site <- c(rep("C2A", 14), rep("C2B", 14), rep("C5A", 14), rep("C5B", 14),
#          rep("D2A", 14), rep("D2B", 14), rep("D5A", 14), rep("D5B", 14))
#print(site)

#rep <- c("R0", rep("R1",5), rep("R2", 4), rep("R3",4))  %>% 
#  rep(8)
#print(rep)

sensor.list <- as_tibble(read_csv("../sensor_list.csv"))
#view(sensor.list)

#Add column to DDsums.df and sensor.list to give a unique column to merge with.
DDsums.df.A <- DDsums.df.pre %>% 
  mutate(srps.name = paste(site, rep, position, season, sep = "_"))
#head(DDsums.df.A)

sensor.list.A <- sensor.list %>% 
  mutate(srp.name = paste(site, rep, position, sep = "_"))
#head(sensor.list.A)

#DDsums.df.new <- merge(x = sensor.list.A, y = DDsums.df.A, by = "srp.name", all.x = TRUE)
DDsums.df <- merge(x = sensor.list.A, y = DDsums.df.A, all.x = TRUE)


view(DDsums.df)



#```

#### Replace m01surf and m02surf labels

```{r remove m01surf and m02surf}

#format:
#df[df == "Old Value"] <- "New Value"

DDsums.df[DDsums.df == "m01surf"] <- "lsurf"
DDsums.df[DDsums.df == "m02surf"] <- "m0surf"

#print(DDsums.df$position == "m01surf" | DDsums.df$position == "m02surf")

```
