---
title: "File_pipeline"
author: "Sara"
date: "11/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(glue)
library(stringr)


```

## List Files

```{r }
m <- list.files("./RAW_DATA/", pattern = "*.csv")

file_df <- data.frame(paths = m, name = str_replace(m, ".csv",""), full_paths = list.files("./RAW_DATA/", pattern = "*.csv", full.names = TRUE))

f <- file_df$name #intermediate step, assigning "name" column to a variable to be used in g.
g <- as.data.frame(str_split_fixed(f, "_", 3))

master <- cbind(file_df,g)
colnames(master) <- c("paths", "name", "full_paths", "sensor", "serial" , "upload_date")


head(master)
```
## Endgame:
We will be using pwalk or map2 from tidyverse to iterate the create_csv function over a list of files. We will first create the function, and then we will create the list of inputs to use with our iterating function (pwalk or map2).

## Sara's function

-function, requiring the inputs full_paths and sensor. 
-create object "temps" by reading each csv file, skip the first 20 rows.
-column names within temps will be: "date.time", "unit", "value". Change these titles as needed.
-write a new csv using the object temps. File path will be: "./CLEAN_DATA{sensor}_2020.csv"

"./" means in working directory

## Creating the input list
create new dataframe: "df_master" by putting master through the following:
select the full_path and sensor columns, mutate the full_paths column to character.
  -we did this because R was giving us an error that full_paths needed to be a string.

Last line of code: create_csv function test on iButton 11.
It's a good idea to run a test on one file before using pwalk or map2, which will run the function over the whole set.

## Iterate
Use pwalk or map2. In this case, pwalk is fed the dataframe df_master, and will perform the function create_csv over each row.
map2 uses 2 columns (df_master$full_paths, df_master$sensor) and uses the row position to match the inputs together. Both approaches will work.

```{r }

create_csv <- function(full_paths, sensor){
  temps <- read_csv(full_paths, skip = 20)
  colnames(temps) <- c("date.time", "unit", "value")
  write_csv(temps, glue("./CLEAN_DATA/{sensor}_2020.csv"))
}

#Test the function
create_csv("./RAW_DATA/i11_3E00000069B0E841_083020.csv", "i11")


#creating list of full_paths and sensors for input into the create_csv function
df_master <- master %>% 
  select(full_paths, sensor) %>% 
  mutate(full_paths = as.character(full_paths))

head(df_master)


#Iterate! If you want to see something really satisfying, open the folder that you're saving these to and watch the files pop in.

pwalk(df_master, create_csv)


map2(df_master$full_paths, df_master$sensor, create_csv)


```


##Let's try this on the remaining files.
changed location name to RAW_DATA_B2

```{r }


m <- list.files("./RAW_DATA_B2/", pattern = "*.csv")
m

file_df <- data.frame(paths = m, name = str_replace(m, ".csv",""), full_paths = list.files("./RAW_DATA_B2/", pattern = "*.csv", full.names = TRUE))

summary(file_df)

f <- file_df$name 
g <- as.data.frame(str_split_fixed(f, "_", 3))


master <- cbind(file_df,g)
colnames(master) <- c("paths", "name", "full_paths", "sensor", "serial" , "upload_date")

summary(master)

create_csv <- function(full_paths, sensor){
  temps <- read_csv(full_paths, skip = 20)
  colnames(temps) <- c("date.time", "unit", "value")
  write_csv(temps, glue("./CLEAN_DATA/{sensor}_2020.csv"))
}

df_master <- master %>% 
  select(full_paths, sensor) %>% 
  mutate(full_paths = as.character(full_paths))

summary(df_master)

pwalk(df_master, create_csv)


map2(df_master$full_paths, df_master$sensor, create_csv)


create_csv("./RAW_DATA_B2/i11_3E00000069B0E841_083020.csv", "i11")



```
## Not sure why there are missing column names filled in. In the newly-created csv file, 3-4 columns are filled with NA. 
What's going on here?
