---
title: "Itasca linear mixed-effects models"
author: "Hannah Friesen and Sara DeLaurentis"
date: "1/8/2022"
output: html_document
---


I'm not sure what this first section does...
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(stringr)
library(stringi)
library(dplyr)
library(ggplot2)
library(lubridate)



```

## R Markdown
#Step 1: Create an object

```{r }

WholeSeason_19.20.lme <- lme(weeklyAvg~Treatment*week.x*Sensor.depth,
                       random=list(Site = pdDiag(~week.x*Treatment)),
                       #Sensor.depth=~weekBeginDate),
                       correlation=corAR1(form=~weekBeginDate|Site/Treatment/Sensor.depth),
                       data=WholeSeason_19.20,
                       na.action=na.exclude) 


```
WholeSeason_19.20.lme = the new R object containing the output of the mixed effect model
weeklyAvg = a continuous variable containing temperature averaged over weekly timescales
Treatment = a factor variable containing information about what treatment was applied (e.g., "Removal" or "Control", in this case)
week.x = a factor variable created from weekBeginDate (a Date variable) (e.g., 2017-10-21 date variable = 1 week.x factor variable, etc.)
Sensor.depth = a continuous variable describing the depth below the soil at which a sensor was installed
Site = a factor variable describing the research site location
weekBeginDate = a date variable describing the date beginning the week for the weekly averaged temperature data (weeklyAvg)
WholeSeason_19.20 = the R object containing the data being used to generate the mixed effect model


##Highlight points where treatments were significantly different

And here is the code for the comparison of regression lines to highlight points where the treatments were significantly different from one another.:

```{r }
#Early Winter 2017-2018
WholeSeason_19.20.lsm <- lsmeans(WholeSeason_19.20.lme, ~Treatment | week.x | Sensor.depth)
summary(WholeSeason_19.20.lsm)
#where WholeSeason_19.20.lsm is the least square means (I think this is now emmeans - should work the same) object of the mixed effects model WholeSeason_19.20.lme, with Treatment as a function of week.x as a function of Sensor.depth (variables described below)

#CONTRAST STATEMENTS
EW1718.pairwise <- contrast(WholeSeason_19.20.lsm, alpha=0.05, method="revpairwise", by=c("Sensor.depth","week.x"))
EW1718.pairwise
#where EW1718.pairwise is the regression line comparison of the least square means (WholeSeason_19.20.lsm) by Sensor.depth and week.x

 

```
WholeSeason_19.20.lme - mixed effects model (described above)
WholeSeason_19.20.lsm - least square means object
EW1718.pairwise - pairwise regression line comparison by Sensor.depth and week.x
Treatment - factor variable describing the treatment type; in this case, "Removal" or "Control"
Sensor.depth - continuous variable describing the soil depth at which sensors were installed (e.g. 5, 10, 15, 20, 40, and 60cm)
week.x - factor variable created from weekBeginDate (a Date variable) (e.g., 2017-10-21 date variable = 1 week.x factor variable, etc.)



#Summarizing data by week:
Set date format.
Create object week.START to start the point of reference for counting the weeks
Create a column "temps$week.begin", using floor_date to calculate the start date of each week.
Create a column, temps$week.no, that subtracts the current date from the week.START and adds 1.
Then group the data by week.no and calculate the weekly mean.

```{r }
temps <- read.csv("C2A_R3_m0surf_i39_2020.csv")
view(temps)
class(temps$longdate)

temps$longdate <- as.Date(temps$longdate) 
week.START <- as.Date("2019-11-10")
temps$week.begin <- floor_date(as.Date(temps$longdate), unit = "weeks")
temps$week.no <- 1 + difftime(temps$week.begin, week.START, units = "weeks")

  
temps <- temps %>%  
  group_by(week.no) %>%
  mutate(week_avg = mean(value)) %>%  #mutate will create another column, but will not condense to a single value per week. To condense, which I think is optimal, use summarise/summarize.
  summarize
  
view(temps)


  
```


```{r }

```
