---
title: "Thermal time"
author: "Sara DeLaurentis"
date: "2/1/2022"
output: html_document
purpose: tabulate (and plot?) thermal time for Itasca Project
---


Archived 2024-05-24
Kept for potential usefulness with week calculations.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(stringi)
library(stringr)
library(dplyr)
library(ggplot2)
library(purrr)
library(lubridate)

library(nlme)
library(car)



```

## R Markdown


```{r }

#for one file testing
temps <- read.csv("C2A_R3_m0surf_i39_2020.csv")
head(temps)
temps$longdate <- as.POSIXct(temps$longdate)


#for multiple file testing (4 or more files merged together)

csv_files <- list.files(pattern = ".csv", full.names = TRUE)
head(csv_files)

merged_ibuttons <- map_dfr(csv_files, read.csv) #map_dfr reads csv files and joins them by rowbind.
head(merged_ibuttons)
str(merged_ibuttons)

temps <- merged_ibuttons

#write.csv(merged_ibuttons, file = "merged_ibuttons.csv")


str(temps)

View(temps)



temps <- temps %>% 
  group_by(labels, longdate) %>% #groups by label and longdate. this keeps all the ibuttons separate during this process.
  mutate(daily_avg = mean(value)) %>%  #create daily_avg column
  distinct(longdate, .keep_all = TRUE) %>% #only keep distinct longdate rows (for each ibutton label group)
  ungroup() 
head(temps)



plot(daily_avg ~ longdate, data = temps)
#developing thermal time code section

#sum(df$column == value, na.rm=TRUE)

sum(temps$value >= 0, na.rm=TRUE)



```

##Linear model practice


```{r }


fm1 <- lm(daily_avg ~ longdate, data = temps) #daily avg as a function of date
summary(fm1) 
plot(daily_avg ~ longdate, data = temps)
lines(fitted.values(fm1))
residualPlot(fm1)

durbinWatsonTest(fm1)



fm2 <- lm(daily_avg ~ treatment * as.Date(longdate), data = temps)

#data set as of 2/2/2022, 10pm. All the same depth, sites C2B and C5A, R1 and R2 at each: 
#"./C2B_R1_m0surf_i105_2020.csv" 
#"./C2B_R2_m0surf_i101_2020.csv" 
#"./C5A_R1_m0surf_i16_2020.csv"  
#"./C5A_R2_m0surf_i21_2020.csv"

summary(fm2)
anova(fm2)

residualPlot(fm2)

durbinWatsonTest(fm2)


fm3 <- lme(daily_avg ~ treatment * julian(as.Date(longdate)), data = temps, random = ~ 1 | treatment)
#NOTE: julian calculates the number of days since the origin. The default origin is 1970-01-01.

summary(fm3)
#residualPlot(fm3)
durbinWatsonTest(fm3)

AIC(fm3)

```
##Weekly summaries
2/3/2022: It's a CLEAN CHUNK
Requires: 
  -one or more csv files saved in the current directory.
  -a column with dates only, labeled "longdate"
  -metadata "labels" column attached, so the data can be grouped by them

```{r }

#for multiple file testing (4 or more files merged together)

csv_files <- list.files(pattern = ".csv", full.names = TRUE)
head(csv_files)


#data set as of 2/2/2022, 10pm. All the same depth, sites C2B and C5A, R1 and R2 at each: 
#"./C2B_R1_m0surf_i105_2020.csv" 
#"./C2B_R2_m0surf_i101_2020.csv" 
#"./C5A_R1_m0surf_i16_2020.csv"  
#"./C5A_R2_m0surf_i21_2020.csv"


merged_ibuttons <- map_dfr(csv_files, read.csv) #map_dfr reads csv files and joins them by rowbind.
head(merged_ibuttons)
str(merged_ibuttons)

temps <- merged_ibuttons

#write.csv(merged_ibuttons, file = "merged_ibuttons.csv") #if you want to create a csv from the four merged files.

str(temps)
View(temps)


#Summarizing data by week:
#Set date format.
#Create object week.START to start the point of reference for counting the weeks
#Create a column "temps$week.begin", using floor_date to calculate the start date of each week.
#Create a column, temps$week.no, that subtracts the current date from the week.START and adds 1.
#Then group the data by week.no and calculate the weekly mean.

#```{r }
#for one file practice:
#temps <- read.csv("C2A_R3_m0surf_i39_2020.csv")
#view(temps)
#class(temps$longdate)
#class(week.START)


## for these tests, START HERE ##
temps <- merged_ibuttons
temps$longdate <- as.Date(temps$longdate) 
week.START <- as.Date("2019-11-10") #assign variable for week.START
week.END <- as.Date("2020-06-30")
temps <- temps %>% 
  filter(longdate >= as.Date(week.START) & longdate <= as.Date(week.END))
temps$week.begin <- floor_date(as.Date(temps$longdate), unit = "weeks") #create column of 1st day of each week
temps$week.no <- 1 + difftime(temps$week.begin, week.START, units = "weeks") #week number, starting at origin. Instead of a column in a dataframe, it could potentially be part of a formula.
  
temps <- temps %>%  
  group_by(labels, week.no) %>% #VERY IMPORTANT to include labels in the grouping structure.
  mutate(week_avg = mean(value)) %>%  #mutate will create another column, but will not condense to a single value per week. 
  distinct(week.no, .keep_all = TRUE) %>% #only keep distinct week.no rows (for each ibutton label group)
  ungroup()
temps$value = NULL
head(temps)
view(temps)
str(temps)


#ggplot(week_avg ~ week.begin, data = temps, type = "p", col = c("red", "orange", "blue", "purple"))

I = ggplot() +
  geom_line(data = temps, aes(x = as.Date(week.begin), y = week_avg, group = labels, colour = labels)) +
  ggtitle("Quick plot: Weekly average vs. Week start date", subtitle = "Mineral surface sensors at C2B and C5A")
print(I)

```
